### 2020-09-17 paper/cover-letter/cover-letter.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
% It was modified by Katy Huff
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%    PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[10.5pt]{letter} % Default font size of the document, change to 10pt to fit more text
\usepackage{graphicx}
%\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font

% Margins
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
%\let\raggedleft\raggedright % Pushes the date (at the top) to the left, comment this line to have the date on the right

\usepackage{eso-pic,graphicx}

%--------------------------------------------------------------------------------------
%--------INPUT DATA
%--------------------------------------------------------------------------------------
\usepackage{xspace}
\newcommand{\StudentFirstName}{O.\xspace}
\newcommand{\StudentLastName}{Ashraf\xspace}
\newcommand{\RecipientName}{Prof. Mostafa Ghiaasiaan\xspace}
\newcommand{\RecipientAddress}{Executive Editor\\Annals of Nuclear Energy}
%\newcommand{\<++>}{<++>\xspace}

\begin{document}
	\AddToShipoutPictureBG*{\includegraphics[width=\paperwidth,height=\paperheight]{background.pdf}}

	%----------------------------------------------------------------------------------------
	%    ADDRESSEE SECTION
	%----------------------------------------------------------------------------------------

	\begin{letter}{\RecipientName\\
			\RecipientAddress\xspace}

		%-------------------------------------------------------------------------------
		%    YOUR NAME & ADDRESS SECTION
		%-------------------------------------------------------------------------------
		\address{O. Ashraf\\
			oabdelaziz@mephi.ru\\
			osama.ashraf@edu.asu.edu.eg\\
			Institute of Nuc. Physics and Eng.\\
			National Research Nuclear University\\
			Moscow, Russia, 115409}

		%-------------------------------------------------------------------------------
		%    LETTER CONTENT SECTION
		%-------------------------------------------------------------------------------

		\opening{To \RecipientName,}

		Please consider our revised manuscript, ``Preliminary design of control rods in the single-fluid double-zone thorium molten salt reactor (SD-TMSR)'' which I and my coauthors (Dr. Andrei Rykhlevskii, Prof. G. V. Tikhomirov, and Prof. Kathryn D. Huff) are submitting for publication as a research article 
		in Annals of Nuclear Energy.

		We appreciate the interest that the editors and reviewers have taken in our manuscript and the constructive suggestions they have given. We have addressed the major concerns of the reviewers. It is our belief that the manuscript is substantially improved after making the suggested edits.

		Following this letter is the detailed response to reviewers. We included the 
		answers in the manuscript. Changes made in the manuscript are marked using track changes.

	Thank you again for consideration of our revised manuscript.

		%Declarations of interest: none\\

		\closing{Sincere regards,
	\includegraphics[height=1.5cm]{signature.png}\\
	\fromsig{O. Ashraf\\
		Assistant Teacher\\
		Institute of Nuc. Physics and Eng.\\
		National Research Nuclear University (MEPhI)}
}

%-------------------------------------------------------------------------------

\end{letter}

\end{document}